---
title: "Miscellaneous Interactive Visuals"
author: "Xiaotang (Jeffrey) Zhou"
output: 
    html_document:
        toc: TRUE
        toc_float: TRUE
---

<br>

These are additional interactive visuals that are derived from the information on all of the goals ever scored by Ovechkin and Gretzky. Although not relevant to the scope of the project, these visuals still provide a comparison between the individual goal scoring talents of both players.

Note that as Ovechkin is currently an active player in the NHL, any data corresponding to him is accurate as of the last update to this website. This update was most recently performed on December 15th 2025. 

<br>

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}

library(data.table)
library(tidyverse)
library(dplyr)
library(plotly)
library(DT)
library(knitr)
library(RCurl)

# Initialize code chunk options
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = TRUE,
  cache = FALSE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px",
  class.source = "code-r")

ovi_data <- getURL("https://raw.githubusercontent.com/jeffreyz374/JSC370_Project/main/Data/ovi-data.csv")
ovi_data <- read.csv(text = ovi_data)
  

gretzky_data <- getURL("https://raw.githubusercontent.com/jeffreyz374/JSC370_Project/main/Data/gretzky-data.csv")
gretzky_data <- read.csv(text = gretzky_data)

ovi_goals_by_date <- ovi_data %>%
  group_by(Date) %>%
  summarise(Goals = n())

unique_date_opponent_pairs <- ovi_data %>%
  select(c("Date", "Opponent")) %>%
  distinct(Date, Opponent)

ovi_goals_by_date <- merge(ovi_goals_by_date, unique_date_opponent_pairs)

caps_wins_by_date <- ovi_data %>%
  distinct(Date, Result)

x_ovi <- merge(ovi_goals_by_date, caps_wins_by_date)

gretzky_goals_by_date <- gretzky_data %>%
  group_by(Date) %>%
  summarise(Goals = n())

wins_by_date <- gretzky_data %>%
  distinct(Date, Result)

x_gretzky <- merge(gretzky_goals_by_date, wins_by_date)

ovi_potential_games <- getURL("https://raw.githubusercontent.com/jeffreyz374/JSC370_Project/main/Data/ovi-potential-games.csv")
ovi_potential_games <- read.csv(text = ovi_potential_games)

gretzky_potential_games <- getURL("https://raw.githubusercontent.com/jeffreyz374/JSC370_Project/main/Data/gretzky-potential-games.csv")
gretzky_potential_games <- read.csv(text = gretzky_potential_games)

name_table <- getURL("https://raw.githubusercontent.com/jeffreyz374/JSC370_Project/main/Data/name-table.csv")
name_table <- read.csv(text = name_table)

```

```{r, echo = FALSE}
ovi_data_copy <- ovi_data
ovi_data_copy[is.na(ovi_data_copy)] <- "N/A"
ovi_data_copy$PS <- case_when(ovi_data_copy$PS == 0 ~ "No", ovi_data_copy$PS == 1 ~ "Yes")
ovi_data_copy$ENG <- case_when(ovi_data_copy$ENG == 0 ~ "No", ovi_data_copy$ENG == 1 ~ "Yes")
ovi_data_copy$Result <- case_when(ovi_data_copy$Result == "W" ~ "Regulation Win",
                                  ovi_data_copy$Result == "L" ~ "Regulation Loss",
                                  ovi_data_copy$Result == "OTW" ~ "Overtime Win",
                                  ovi_data_copy$Result == "OTL" ~ "Overtime Loss",
                                  ovi_data_copy$Result == "SOW" ~ "Shootout Win",
                                  ovi_data_copy$Result == "SOL" ~ "Shootout Loss")
ovi_data_copy$HomeAway <- case_when(ovi_data_copy$HomeAway == "H" ~ "Home", ovi_data_copy$HomeAway == "A" ~ "Away")
ovi_data_copy$Strength <- case_when(ovi_data_copy$Strength == "EV" ~ "Even Strength", ovi_data_copy$Strength == "PP" ~ "Power Play", ovi_data_copy$Strength == "SH" ~ "Shorthanded")
ovi_data_copy$Time <- substr(ovi_data_copy$Time, start = 1, stop = 5)
ovi_data_copy$Period <- case_when(ovi_data_copy$Period == 1 ~ "1st", ovi_data_copy$Period == 2 ~ "2nd", ovi_data_copy$Period == 3 ~ "3rd", ovi_data_copy$Period == "OT" ~ "Overtime")
ovi_data_copy$Team <- rep("Washington Capitals", nrow(ovi_data_copy))

ovi_data_copy <- merge(ovi_data_copy, name_table, by.x = c("Opponent"), by.y = c("Abbreviation"))

ovi_data_copy <- ovi_data_copy[c(2:ncol(ovi_data_copy))]

colnames(ovi_data_copy) <- c("Date", "Team", "Age (years)", "1st Assist", "2nd Assist", "Goalie", "Period", "Time", "Strength", "Arena", "Home/Away", "Result", "Empty Net Goal", "Penalty Shot Goal", "Opponent")

ovi_data_copy <- ovi_data_copy[, c(1, 2, 15, 3:14)]

ovi_data_copy <- left_join(ovi_data_copy, ovi_potential_games[c(1, 3, 4)], by = "Date") %>%
  mutate(Final = paste(GF, "-", GA))

ovi_data_copy <- ovi_data_copy[, c(1:13, 18, 14:15)]

colnames(ovi_data_copy) <- c("Date", "Team", "Opponent", "Age (years)", "1st Assist", "2nd Assist", "Goalie", "Period", "Time", "Strength", "Arena", "Home/Away", "Result", "Final Score", "Empty Net Goal", "Penalty Shot Goal")

ovi_data_copy <- ovi_data_copy %>%
  arrange(desc(Date), desc(Period), desc(Time))

ovi_data_copy$cgoal <- rev(c(1:nrow(ovi_data_copy)))

ovi_data_copy <- ovi_data_copy[, c(1:4, 17, 5:16)]

colnames(ovi_data_copy) <- c("Date", "Team", "Opponent", "Age (years)", "Career Goal Number", "1st Assist", "2nd Assist", "Goalie", "Period", "Time", "Strength", "Arena", "Home/Away", "Result", "Final Score", "Empty Net Goal", "Penalty Shot Goal")
```

```{r, echo = FALSE}
gretzky_data_copy <- gretzky_data
gretzky_data_copy[is.na(gretzky_data_copy)] <- "N/A"
gretzky_data_copy$PS <- case_when(gretzky_data_copy$PS == 0 ~ "No", gretzky_data_copy$PS == 1 ~ "Yes")
gretzky_data_copy$ENG <- case_when(gretzky_data_copy$ENG == 0 ~ "No", gretzky_data_copy$ENG == 1 ~ "Yes")
gretzky_data_copy$Result <- case_when(gretzky_data_copy$Result == "W" ~ "Regulation Win",
                                      gretzky_data_copy$Result == "L" ~ "Regulation Loss",
                                      gretzky_data_copy$Result == "OTW" ~ "Overtime Win",
                                      gretzky_data_copy$Result == "OTL" ~ "Overtime Loss",
                                      gretzky_data_copy$Result == "T" ~ "Tie")
gretzky_data_copy$HomeAway <- case_when(gretzky_data_copy$HomeAway == "H" ~ "Home", gretzky_data_copy$HomeAway == "A" ~ "Away")
gretzky_data_copy$Strength <- case_when(gretzky_data_copy$Strength == "EV" ~ "Even Strength", gretzky_data_copy$Strength == "PP" ~ "Power Play", gretzky_data_copy$Strength == "SH" ~ "Shorthanded")
gretzky_data_copy$Time <- substr(gretzky_data_copy$Time, start = 12, stop = 16)
gretzky_data_copy$Period <- case_when(gretzky_data_copy$Period == 1 ~ "1st", gretzky_data_copy$Period == 2 ~ "2nd", gretzky_data_copy$Period == 3 ~ "3rd", gretzky_data_copy$Period == "OT" ~ "Overtime")
gretzky_data_copy$Team <- case_when(gretzky_data_copy$Team == "EDM" ~ "Edmonton Oilers",
                                    gretzky_data_copy$Team == "LAK" ~ "Los Angeles Kings",
                                    gretzky_data_copy$Team == "STL" ~ "St. Louis Blues",
                                    gretzky_data_copy$Team == "NYR" ~ "New York Rangers")

gretzky_data_copy <- merge(gretzky_data_copy, name_table, by.x = c("Opponent"), by.y = c("Abbreviation"))

gretzky_data_copy <- gretzky_data_copy[c(2:ncol(gretzky_data_copy))]

colnames(gretzky_data_copy) <- c("Date", "Team", "Age (years)", "1st Assist", "2nd Assist", "Goalie", "Period", "Time", "Strength", "Arena", "Result", "Home/Away", "Empty Net Goal", "Penalty Shot Goal", "Opponent")

gretzky_data_copy <- gretzky_data_copy[, c(1, 2, 15, 3:10, 12, 11, 13, 14)]

gretzky_data_copy <- left_join(gretzky_data_copy, gretzky_potential_games[c(1, 3, 4)], by = "Date") %>%
  mutate(Final = paste(GF, "-", GA))

gretzky_data_copy <- gretzky_data_copy[, c(1:13, 18, 14:15)]

colnames(gretzky_data_copy) <- c("Date", "Team", "Opponent", "Age (years)", "1st Assist", "2nd Assist", "Goalie", "Period", "Time", "Strength", "Arena", "Home/Away", "Result", "Final Score", "Empty Net Goal", "Penalty Shot Goal")

gretzky_data_copy <- gretzky_data_copy %>%
  arrange(desc(Date), desc(Period), desc(Time))

gretzky_data_copy$cgoal <- rev(c(1:894))

gretzky_data_copy <- gretzky_data_copy[, c(1:4, 17, 5:16)]

colnames(gretzky_data_copy) <- c("Date", "Team", "Opponent", "Age (years)", "Career Goal Number", "1st Assist", "2nd Assist", "Goalie", "Period", "Time", "Strength", "Arena", "Home/Away", "Result", "Final Score", "Empty Net Goal", "Penalty Shot Goal")
```

```{css, echo = FALSE}
.code-r { /* Code block */
  font-size: 15px;
}

.code-r-small { /* Code block */
  font-size: 10px;
}
```

<br>

## Ovechkin and Gretzky Goals By Season {.tabset}

This interactive visual shows the number of goals scored by Ovechkin and Gretzky grouped by the NHL season in which each player scored those goals. 

Note that as Ovechkin is currently an active player in the NHL, any data corresponding to him is accurate as of the last update to this website. This update was most recently performed on December 15th 2025.

```{r, echo = FALSE}
ovi_season_stats <- ovi_data %>%
  mutate(Season = case_when(Date >= "2005-10-01" & Date <= "2006-06-01" ~ "2005-06",
                            Date >= "2006-10-01" & Date <= "2007-06-01" ~ "2006-07",
                            Date >= "2007-10-01" & Date <= "2008-06-01" ~ "2007-08",
                            Date >= "2008-10-01" & Date <= "2009-06-01" ~ "2008-09",
                            Date >= "2009-10-01" & Date <= "2010-06-01" ~ "2009-10",
                            Date >= "2010-10-01" & Date <= "2011-06-01" ~ "2010-11",
                            Date >= "2011-10-01" & Date <= "2012-06-01" ~ "2011-12",
                            Date >= "2012-10-01" & Date <= "2013-06-01" ~ "2012-13",
                            Date >= "2013-10-01" & Date <= "2014-06-01" ~ "2013-14",
                            Date >= "2014-10-01" & Date <= "2015-06-01" ~ "2014-15",
                            Date >= "2015-10-01" & Date <= "2016-06-01" ~ "2015-16",
                            Date >= "2016-10-01" & Date <= "2017-06-01" ~ "2016-17",
                            Date >= "2017-10-01" & Date <= "2018-06-01" ~ "2017-18",
                            Date >= "2018-10-01" & Date <= "2019-06-01" ~ "2018-19",
                            Date >= "2019-10-01" & Date <= "2020-06-01" ~ "2019-20",
                            Date >= "2020-10-01" & Date <= "2021-06-01" ~ "2020-21",
                            Date >= "2021-10-01" & Date <= "2022-06-01" ~ "2021-22",
                            Date >= "2022-10-01" & Date <= "2023-06-01" ~ "2022-23",
                            Date >= "2023-10-01" & Date <= "2024-06-01" ~ "2023-24",
                            Date >= "2024-10-01" & Date <= "2025-06-01" ~ "2024-25",
                            Date >= "2025-10-01" & Date <= "2026-06-01" ~ "2025-26"))

ovi_season_goals <- ovi_season_stats %>%
  group_by(Season) %>%
  summarise(Goals = n()) %>%
  ggplot(aes(x = Season, Goals)) +
  geom_bar(stat = "identity", fill = "red") +
  ylab("Goals") +
  ggtitle("Number of goals scored by Ovechkin at varying seasons in his NHL career") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

# ovi_season_engs <- ovi_season_stats %>%
#   group_by(Season) %>%
#   summarise(ENGs = sum(ENG),
#             NonENGs = n() - ENGs) %>%
#   pivot_longer(-Season) %>%
#   ggplot(aes(x = Season, y = value, fill = name)) +
#   geom_col(position = position_stack()) +
#   theme_minimal() +
#   theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

gretzky_season_stats <- gretzky_data %>%
  mutate(Season = case_when(Date >= "1979-10-01" & Date <= "1980-06-01" ~ "1979-80",
                            Date >= "1980-10-01" & Date <= "1981-06-01" ~ "1980-81",
                            Date >= "1981-10-01" & Date <= "1982-06-01" ~ "1981-82",
                            Date >= "1982-10-01" & Date <= "1983-06-01" ~ "1982-83",
                            Date >= "1983-10-01" & Date <= "1984-06-01" ~ "1983-84",
                            Date >= "1984-10-01" & Date <= "1985-06-01" ~ "1984-85",
                            Date >= "1985-10-01" & Date <= "1986-06-01" ~ "1985-86",
                            Date >= "1986-10-01" & Date <= "1987-06-01" ~ "1986-87",
                            Date >= "1987-10-01" & Date <= "1988-06-01" ~ "1987-88",
                            Date >= "1988-10-01" & Date <= "1989-06-01" ~ "1988-89",
                            Date >= "1989-10-01" & Date <= "1990-06-01" ~ "1989-90",
                            Date >= "1990-10-01" & Date <= "1991-06-01" ~ "1990-91",
                            Date >= "1991-10-01" & Date <= "1992-06-01" ~ "1991-92",
                            Date >= "1992-10-01" & Date <= "1993-06-01" ~ "1992-93",
                            Date >= "1993-10-01" & Date <= "1994-06-01" ~ "1993-94",
                            Date >= "1994-10-01" & Date <= "1995-06-01" ~ "1994-95",
                            Date >= "1995-10-01" & Date <= "1996-06-01" ~ "1995-96",
                            Date >= "1996-10-01" & Date <= "1997-06-01" ~ "1996-97",
                            Date >= "1997-10-01" & Date <= "1998-06-01" ~ "1997-98",
                            Date >= "1998-10-01" & Date <= "1999-06-01" ~ "1998-99"))

gretzky_season_goals <- gretzky_season_stats %>%
  group_by(Season) %>%
  summarise(Goals = n()) %>%
  ggplot(aes(x = Season, Goals)) +
  geom_bar(stat = "identity", fill = "blue") +
  ylab("Goals") +
  ggtitle("Number of goals scored by Gretzky at varying seasons in his NHL career") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

# gretzky_season_engs <- gretzky_season_stats %>%
#   group_by(Season) %>%
#   summarise(ENGs = sum(ENG),
#             NonENGs = n() - ENGs) %>%
#   pivot_longer(-Season) %>%
#   ggplot(aes(x = Season, y = value, fill = name)) +
#   geom_col(position = position_stack()) +
#   theme_minimal() +
#   theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```

### Ovechkin

```{r, echo = FALSE}
ggplotly(ovi_season_goals) %>%
  layout(hoverlabel=list(bgcolor="white"))
```

### Gretzky

```{r, echo = FALSE}
ggplotly(gretzky_season_goals) %>%
  layout(hoverlabel=list(bgcolor="white"))
```
## {-}

<br>

## Ovechkin and Gretzky Goals By Age {.tabset}

This interactive visual shows the number of goals scored by Ovechkin and Gretzky grouped by the age at which each player scored those goals.

As Ovechkin was born on September 17th 1985, the visual corresponding to him will be identical to the previous one depicting the frequency of his goals grouped by NHL season, since September is not within the NHL season.

In contrast, as Gretzky was born on January 26th 1961, the visual corresponding to him is different from the previous one depicting the frequency of his goals grouped by NHL season, since January is in the middle of the NHL season.

Note that as Ovechkin is currently an active player in the NHL, any data corresponding to him is accurate as of the last update to this website. This update was most recently performed on December 15th 2025.

```{r, echo = FALSE}
gretzky_age_stats <- gretzky_data %>%
  group_by(Age) %>%
  summarise(Goals = n())

ovi_age_stats <- ovi_data %>%
  group_by(Age) %>%
  summarise(Goals = n())

ovi_age_goals <- ovi_age_stats %>%
  ggplot(aes(x = Age, Goals)) +
  geom_bar(stat = "identity", fill = "red") +
  ylab("Goals") +
  ggtitle("Number of goals scored by Ovechkin at varying ages in his NHL career") +
  theme_minimal()

gretzky_age_goals <- gretzky_age_stats %>%
  ggplot(aes(x = Age, Goals)) +
  geom_bar(stat = "identity", fill = "blue") +
  ylab("Goals") +
  ggtitle("Number of goals scored by Gretzky at varying ages in his NHL career") +
  theme_minimal()

age_joint <- left_join(gretzky_age_stats, ovi_age_stats, by = "Age") %>%
  rename(Gretzky = Goals.x, Ovechkin = Goals.y)
age_joint[is.na(age_joint)] <- 0

ovi_gretzky_age_joint <- age_joint %>%
  gather(Player, Goals, -Age) %>%
  ggplot(aes(x = Age, y = Goals, fill = Player)) +
  scale_fill_manual(values = c("blue", "red")) +
  geom_col(position = "dodge") +
  ylab("Goals") +
  ggtitle("Comparison of the number of goals scored by both \n Ovechkin and Gretzky at varying ages in their NHL careers") +
  theme_minimal()
```

### Ovechkin

```{r, echo = FALSE}
ggplotly(ovi_age_goals) %>%
  layout(hoverlabel=list(bgcolor="white"))
```

### Gretzky

```{r, echo = FALSE}
ggplotly(gretzky_age_goals) %>%
  layout(hoverlabel=list(bgcolor="white"))
```

### Both

```{r, echo = FALSE}
ggplotly(ovi_gretzky_age_joint)%>%
  layout(hoverlabel=list(bgcolor="white"))
```
## {-}

<br>

## Ovechkin and Gretzky Goals By Opponent {.tabset}

This interactive visual shows the number of goals scored by Ovechkin and Gretzky grouped by the opponent that was scored on.

Note that as Ovechkin is currently an active player in the NHL, any data corresponding to him is accurate as of the last update to this website. This update was most recently performed on December 15th 2025.

```{r, echo = FALSE}
gretzky_opponent_stats <- gretzky_data_copy %>%
  group_by(Opponent) %>%
  summarise(Goals = n())

ovi_opponent_stats <- ovi_data_copy %>%
  group_by(Opponent) %>%
  summarise(Goals = n())

ovi_opponent_goals <- ovi_opponent_stats %>%
  ggplot(aes(x = reorder(Opponent, Goals), y = Goals, text = paste("Opponent: ", Opponent, '\nGoals: ', Goals))) +
  geom_bar(stat = "identity", fill = "red") +
  ylab("Goals") +
  xlab("Opponent") +
  ggtitle("Number of goals scored by Ovechkin against \n varying opponents in his NHL career") +
  theme_minimal() +
  coord_flip()

gretzky_opponent_goals <- gretzky_opponent_stats %>%
  ggplot(aes(x = reorder(Opponent, Goals), y = Goals, text = paste("Opponent: ", Opponent, '\nGoals: ', Goals))) +
  geom_bar(stat = "identity", fill = "blue") +
  ylab("Goals") +
  xlab("Opponent") +
  ggtitle("Number of goals scored by Gretzky against \n varying opponents in his NHL career") +
  theme_minimal() +
  coord_flip()
```

### Ovechkin

```{r, echo = FALSE}
ggplotly(ovi_opponent_goals, tooltip = c("text")) %>%
  layout(autosize = F, height = 450, hoverlabel=list(bgcolor="white"))
```

### Gretzky

```{r, echo = FALSE}
ggplotly(gretzky_opponent_goals, tooltip = c("text")) %>%
  layout(autosize = F, height = 450, hoverlabel=list(bgcolor="white"))
```

## {-}

<br>

## Ovechkin and Gretzky Goals by Goalie Scored On (Including Empty Net Goals) {.tabset}

This interactive visual shows the number of goals scored by Ovechkin and Gretzky grouped by the goalie on which they scored those goals.

Note that as Ovechkin is currently an active player in the NHL, any data corresponding to him is accurate as of the last update to this website. This update was most recently performed on December 15th 2025.

```{r, echo = FALSE, fig.height = 100}
ovi_goalies <- ovi_data %>%
  mutate(Goalie = replace(Goalie, is.na(Goalie), "Empty Net")) %>%
  group_by(Goalie) %>%
  summarise(Goals = n())

gretzky_goalies <- gretzky_data %>%
  mutate(Goalie = replace(Goalie, is.na(Goalie), "Empty Net")) %>%
  group_by(Goalie) %>%
  summarise(Goals = n())

ovi_goalies_plot <- ovi_goalies %>%
  ggplot(aes(x = reorder(Goalie, Goals), y = Goals, text = paste("Goalie: ", Goalie, '\nGoals: ', Goals))) +
  geom_bar(stat = "identity", fill = "red") +
  ylab("Goals") +
  xlab("Goalie") +
  ggtitle("Number of goals scored on various goalies by \n Ovechkin in his NHL career") +
  theme_minimal() +
  coord_flip()

gretzky_goalies_plot <- gretzky_goalies %>%
  ggplot(aes(x = reorder(Goalie, Goals), y = Goals, text = paste("Goalie: ", Goalie, '\nGoals: ', Goals))) +
  geom_bar(stat = "identity", fill = "blue") +
  ylab("Goals") +
  xlab("Goalie") +
  ggtitle("Number of goals scored on various goalies by \n Gretzky in his NHL career") +
  theme_minimal() +
  coord_flip()
  
```

### Ovechkin

```{r, echo = FALSE, fig.height = 100}
ggplotly(ovi_goalies_plot, tooltip = c("text")) %>%
  layout(autosize = F, height = 2000, hoverlabel=list(bgcolor="white"))
```

### Gretzky

```{r, echo = FALSE, fig.height = 100}
ggplotly(gretzky_goalies_plot, tooltip = c("text")) %>%
  layout(autosize = F, height = 2000, hoverlabel=list(bgcolor="white"))
```

## {-}
